---
title: "EDA.Rmd"
output:
  pdf_document: default
  html_document: default
---

# EDA Autism spectrum discorder quiz
### Wouter Zeevat

\newpage

\tableofcontents
\newpage

## Introduction
The AQ-10 test is a test with 10 questions that will try and predict if you have the autism spectrum disorder. Each question has 4 anwsers. Slightly agree, Agree, Slightly disagree and disagree. Every question gives either 1 or 0 points based on the anwser. Slightly agree gives the same points as agree and the same for disagree. They made it like this to make it feel like slightly agree/disagree feels like a better anwser when in doubt.
The more points the people have, the more chance there is of them having ADS (Autism disorder spectrum).

## Research question
#### How accurate can the AQ-10 test predict whether someone has the autism spectrum disorder?
The goal of this research question is to find out if this autism spectrum disorder test actually works and predicts someone has it.
This would involve machine learning by testing if the computer would find correlations and would be able to predict them actually having ASD, which the test does aswell. This way the accuracy of the test can be measured.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, include=FALSE}
install.packages("ggplot2")  # Install and load ggplot2
install.packages("dplyr")
install.packages("viridis")  # Install
```
```{r, echo=FALSE, include=FALSE}

library("ggplot2")
library("dplyr")
library("viridis") 
```

We will start off by looking at the data and codebook. The data contains 20 variables which will be loaded in as the codebook.
```{r, echo=FALSE}
data <- read.csv('Autism_Data.arff')
codebook <- read.csv('codebook.csv')

colnames(data) <- codebook$Variable.name.short

get_readable <- function(short) {
  if (!short %in% codebook$Variable.name.short) {
    return("This isn't a column")
  } else {
    return(codebook$Variable.name.human.readable[which(codebook$Variable.name.short == short)])
  }
}

knitr::kable(codebook)
```
## The data
This is the data that will be used in the following project. it contains various information about adults doing an autism test. The columns speak for themselves except for the first 10. These columns represent the anwsers of the following question list.

https://www.nice.org.uk/guidance/cg142/resources/autism-spectrum-quotient-aq10-test-pdf-186582493

After knowing all of this it's time to see if the data is right. The data is supposed to have 20 columns and 704 rows. 
```{r, echo=FALSE}
ncol(data)
nrow(data)
```
## Checking the data
The data also needs to be checked of missing data (A row that's missing certain values). The ones that are missing important data will be removed. This needs to be done in order to not mess everything up. For example if someone is missing an anwser of the quiz, their score will be messed up and invalid.

This code will check if there are invalid values in any column.
```{r, echo=FALSE}
data[data == "?"] <- NA
data[1:11] <- data[1:11] %>% mutate_if(is.character,as.numeric) # Changing strings that are numbers to numbers

# Changing the 0,1 to NO and YES. The no means they dont get a point and the yes means they did.
data[1:10][data[1:10] == 0] <- "NO"
data[1:10][data[1:10] == 1] <- "YES"

summary(data)
```
There are two NA's in the data. It is important to remove those in order to keep the data balanced. This will be done by removing their rows.

```{r}
data <- data[-c(which(is.na(data$age))), ]
```

We will now take a look at the ages of the people taking the test are.
```{r}
boxplot(data$age, main="Age of people taking ASD test", ylab="Age", col="cadetblue1")
```
As the boxplot shows, there's one huge outlier. One person would be 383 years old which just isn't humanly possible. The solution to this is taking out the whole row.

```{r, echo=FALSE}
data <- data[-c(which(data$age == 383)),]
boxplot(data$age, main="Age of people taking ASD test", ylab="Age", col="cadetblue1")
boxplot(log10(data$age), main="Age of people taking ASD test", ylab="10log Age", col="cadetblue1")
```

## Correlations

This boxplot shows that there's not much old people (60+) doing the test. The people who take this test are usually mid aged.

Now we will take a look at the test, how much people had what kind of anwser. The goal of this plot is to take a look at what the people scored.
```{r, echo=FALSE}
df <- data.frame(x = rep(paste0("Q", as.character(1:10)), each = 2),
                   y = rep(0, each=20),
                   group = rep(c("YES", "NO"), time = 10))

df$y[1] <- sum(data$a1_score == "YES")
df$y[2] <- sum(data$a1_score == "NO")

df$y[3] <- sum(data$a2_score == "YES")
df$y[4] <- sum(data$a2_score == "NO")

df$y[5] <- sum(data$a3_score == "YES")
df$y[6] <- sum(data$a3_score == "NO")

df$y[7] <- sum(data$a4_score == "YES")
df$y[8] <- sum(data$a4_score == "NO")

df$y[9] <- sum(data$a5_score == "YES")
df$y[10] <- sum(data$a5_score == "NO")

df$y[11] <- sum(data$a6_score == "YES")
df$y[12] <- sum(data$a6_score == "NO")

df$y[13] <- sum(data$a7_score == "YES")
df$y[14] <- sum(data$a7_score == "NO")

df$y[15] <- sum(data$a8_score == "YES")
df$y[16] <- sum(data$a8_score == "NO")

df$y[17] <- sum(data$a9_score == "YES")
df$y[18] <- sum(data$a9_score == "NO")

df$y[19] <- sum(data$a10_score == "YES")
df$y[20] <- sum(data$a10_score == "NO")

ggp <- ggplot(df, aes(x = x, y = y, fill = group, label = y)) +  # Create stacked bar chart
  geom_bar(stat = "identity") + ggtitle("Q10 test Anwsers") + xlab("Question number") + ylab("Amount of anwsers") + scale_fill_manual(values = viridis(5))
ggp 
```

The conclusion of this plot is that most questions are anwsered positively (Without getting a point).

Let's take a look at the correlations now. To start off, the end score will be measured against people actually having ASD. This will give a good view of the test because the test results will directly be compared to them having ASD.

```{r, echo=FALSE}
boxplot(data$end_score ~ data$autism, col=c("cadetblue4", "cadetblue3"), xlab="Has ASD", ylab="Q10 end score", main="Q10 scores vs actually having ASD")
```
As seen in the plot, the scores do actually correlate with someone having ASD. This is true because the scores of the people having ASD are significantly higher than the other people. Let's confirm this by doing a t-test
```{r}
t.test(data$end_score, subset=data$autism, var.equal = TRUE)
```

```{r, echo=FALSE}

df <- data.frame(x = rep(1, each=4),
                   y = rep(0, each=4),
                   group = c("ASD and Family member with ASD", "ASD and no Family member with ASD", "no ASD and Family member with ASD", "no ASD and no Family member with ASD"))

df$y <- c(sum(data$class_asd == "YES" & data$autism == "yes"), sum(data$class_asd == "NO" & data$autism == "yes"),
          sum(data$class_asd == "YES" & data$autism == "no"), sum(data$class_asd == "NO" & data$autism == "no"))

ggp <- ggplot(df, aes(x = x, y = y, fill = group, label = y)) +  # Create stacked bar chart
  geom_bar(stat = "identity") + ggtitle("ASD and family member with ASD") + xlab("") + ylab("amount of people") + scale_fill_manual(values = viridis(5))
ggp
```
This figure shows that Most people that take the test do not have ASD and no family member with ASD. There's also a big part of the group that do not have ASD but have family members with ASD. This makes sense because these people probably want to take the test because they know they have a family member with ASD and are scared they have it too.

```{r, echo=FALSE}
grouped <- data %>% 
  mutate(positives = case_when(
    autism == "yes" ~ "ASD",
    autism == "no" ~ "no ASD"
  ))
ggplot(grouped, aes(x = end_score, y = age, colour = positives)) +
  geom_point() + geom_jitter() + ggtitle("End score compared to age") + geom_smooth(method="lm")
```
Some correlation that would not be good would be the correlation of age and the end score. Something that could be possible is that the older you are, the worse you make the test. This would be bad because then the AQ-10 test would be based on age and not on people actually having ASD. It does not look like this is the case though. Let's confirm this by doing a correlation test!

```{r}
cor.test(data$end_score, data$age, method="pearson")
```
